18.821 - Poker Chips Finite board configurations

### 2014-09-15 FiniteBoards.tex

In this file, we analyze the case where the initial number of chips is finite. That is, the sum of all the chips on the board is finite.

\begin{theorem}
\label{finitestability}
If the initial board has a finite number of chips, there exists an $N$ such that after $N$ earthquakes, the system will be in a stable state (any subsequence earthquake does not cause any chips to move). 
\end{theorem}

@@

\begin{proof}
Let $\Phi: B_f \rightarrow \mathbb{Z}$ be the following function
\[ \Phi(f) = \sum_{x,y} f(x,y)(x+y)^2 \]
Note that this is well-defined since the boards contain a finite number of chips. I will show that 
\[ E(f) \neq f \Rightarrow \Phi(E(f)) > \Phi(f) \]
This means that if there is an earthquake that changes things, then $\Phi(f)$ increases, which means that if $E(f) \neq f$, then for all $N > 0$, $E^N(f) \neq f$, or that you never return to your original state.

We can evaluate $\Phi(E(f)) - \Phi(f)$.
\[ \Phi(E(f)) = \sum_{x,y} E(f(x,y)) (x+y)^2 \]
\[ = \sum_{x,y} \left(f(x,y) + \sum_{l \in n(x,y)} [[f(t) \geq 4]] - 4[[f(x,y) \geq 4]]\right) (x+y)^2 \]
Suppose location $f(x,y) \geq 4$, then the number of chips in location $(x,y)$ will have decreased after applying $E$. For each decrease in of location $(x,y)$ of $4$, there is an increase in two neighbors by $1$.

@@

There exists less than $(x_h - x_l + 2\log S)(y_h - y_l + 2\log S)S$ possible boards $b \in B$ that satisfy this. We know that there are no cycles in the states, so there must be an $0 \geq N < (x_h - x_l + 2\log S)(y_h - y_l + 2\log S)S$ such that $E^N(f) = E^{N+m}(f)$ for all $m \geq 0$.
\end{proof}

>>>>

In this file, we analyze the case where the initial number of chips is finite. That is, the sum of all the chips on the board is finite. 
In this file, we analyze the case where the initial number of chips is finite.
 That is, the sum of all the chips on the board is finite.

\begin{definition}
We say that a board is stable if its values are strictly less than 4. This is equivalent to earthquakes having no effect on the board. 
\end{definition}

\begin{theorem}
\label{finitestability}
If the initial board has a finite number of chips, there exists an $N$ such that after $N$ earthquakes, the resultant board will be in a stable.
\end{theorem}

@@

\begin{proof}
Let $\Phi: B_f \rightarrow \mathbb{Z}$ be the following function
\[ \Phi(f) = \sum_{x,y} f(x,y)(x+y)^2 \]
Note that this is well-defined since the boards contain a finite number of chips. We will show that 
\[ E(f) \neq f \Rightarrow \Phi(E(f)) > \Phi(f) \]
This means that if there is an earthquake that changes things, then $\Phi(f)$ increases, which means that if $E(f) \neq f$, then for all $N > 0$, $E^N(f) \neq f$, or that you never return to your original state.

We can evaluate $\Phi(E(f)) - \Phi(f)$.
\[ \Phi(E(f)) = \sum_{x,y} E(f(x,y)) (x+y)^2 \]
\[ = \sum_{x,y} \left(f(x,y) + \sum_{l \in n(x,y)} [[f(t) \geq 4]] - 4[[f(x,y) \geq 4]]\right) (x+y)^2 \]
Suppose location $f(x,y) \geq 4$, then the number of chips in location $(x,y)$ will have decreased after applying $E$. For each decrease in the number of chips atlocation $(x,y)$ of $4$, there is an increase in its four neighbors neighbors by $1$.

@@

There exists less than $(x_h - x_l + 2\log S)(y_h - y_l + 2\log S)S$ possible boards $b \in B$ that satisfy this. We know that there are no cycles in the states, so there must be an $0 \leq N < (x_h - x_l + 2\log S)(y_h - y_l + 2\log S)S$ such that $E^N(f) = E^{N+m}(f)$ for all $m \geq 0$.
\end{proof}
